---
title: "SuSiE RSS Diagnostic Example"
author: "Yuxin Zou"
date: "2/16/2021"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

Under the null, the posterior distribution of $\hat{z}_j | \hat{z}_{-j}$ is:
$$
\hat{z}_j | \hat{z}_{-j} \sim N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},1/\Omega_{jj}), \Omega = (\sigma^2 R + s^2 I)^{-1}.
$$

To check whether there is an allele switch issue, we test the hypothesis $H_0:$ the sign of $\hat{z}_j$ is correct, vs $H_1:$ the sign of $\hat{z}_j$ is flipped. The likelihood ratio is
$$
LR = \frac{p(-\hat{z}_j | \hat{z}_{-j})}{p(\hat{z}_j | \hat{z}_{-j})}
$$

From the simulation, the standardized differences between observed z score and the conditional mean is longer-tailed than N(0,1). We use a mixture of normals to model the heavier tail empirically. We model the conditional distribution as
$$
\hat{z}_j | \hat{z}_{-j} \sim \sum_{k=1}^{K} \pi_k N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},\frac{\sigma_k^2}{\Omega_{jj}}).
$$
The $\sigma_k$ is a fixed dense grid, the minimum value is $0.8$, the maximum value is $2\sqrt{max(\text{standardized difference}^2)}$. We estimate $\pi_k$ using `mixsqp`. The likelihood ratio test is based on the mixture distribution.


```{r}
dat = readRDS('output/rss_diagnostic/data/data_24.rds')
z = readRDS('output/rss_diagnostic/flip_z/data_24_sim_gaussian_1_get_sumstats_1_flip_z_1.rds')$z
res.GBR = readRDS('output/rss_diagnostic/diagnostic/data_24_sim_gaussian_null_1_get_sumstats_1_flip_z_1_susie_rss_1_diagnostic_1.rds')$res1_simple
res.CLM = readRDS('output/rss_diagnostic/diagnostic/data_24_sim_gaussian_null_1_get_sumstats_1_flip_z_1_susie_rss_3_diagnostic_1.rds')$res1_simple
ld.CLM = dat$ld[['CLM']]
ld.GBR = dat$ld[['sample']]

postmean.GBR = res.GBR$post$postmean
postz.GBR = res.GBR$post$post_z
loglrmix.GBR = as.numeric(res.GBR$post$logLRmix)

postmean.CLM = res.CLM$post$postmean
postz.CLM = res.CLM$post$post_z
logl0.CLM = res.CLM$post$logl0
logl1.CLM = res.CLM$post$logl1
loglr.CLM = res.CLM$post$logLR

logl0mix.CLM = as.numeric(res.CLM$post$logl0mix)
logl1mix.CLM = as.numeric(res.CLM$post$logl1mix)
loglrmix.CLM = as.numeric(res.CLM$post$logLRmix)
```

There are 595 SNPs in this region. The data is simulated under null. The red point (id: 386) in the plot has large LR from the diagnostic.
```{r}
plot(z, pch=20)
points(386, z[386], col='red', pch=20)
```

The estimated $s^2$ using CLM LD is `r res.CLM$s2`.

The QQ plot for the standardized difference:
```{r}
qqnorm(postz.CLM)
qqline(postz.CLM)
```
```{r}
a_min = 0.8
if(max(postz.CLM^2) < 1){
  a_max = 2
}else{
  a_max = 2*sqrt(max(postz.CLM^2))
}
npoint = ceiling(log2(a_max/a_min)/log2(1.05))
a_grid = 1.05^((-npoint):0) * a_max
gd = cbind(res.CLM$post$w[res.CLM$post$w>0], a_grid[res.CLM$post$w>0])
colnames(gd) = c('pi_k', 'sigma_k')
gd
```

The log LR for SNP 386 using mixture model is `r loglrmix.CLM[386]`. 

```{r}
data.frame(observed_z = z[386], conditional_mean_CLM=postmean.CLM[386], conditional_mean_GBR = postmean.GBR[386])
```
The conditional mean from CLM is positive. We check the LD matrix.

```{r}
plot(ld.GBR[386,-386], ld.CLM[386,-386], pch=20, xlab='GBR', ylab='CLM', main='correlations with SNP 386')
abline(v=0,h=0,lty=4)
```











