---
title: "SuSiE RSS Diagnostic using UKB (simulation without signal)"
author: "Yuxin Zou"
date: "3/4/2021"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

We want to

1. Check the consistency between z scores and LD matrix using $s$;

2. Detect allele switch issue (flipped sign in z score) using conditional distribution.

In diagnostic, we assume the following model,
$$
\hat{z} \sim N(0, (1-s) R + s I), 0 \leq s \leq 1.
$$

The posterior distribution of $\hat{z}_j | \hat{z}_{-j}$ is:
$$
\hat{z}_j | \hat{z}_{-j} \sim N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},1/\Omega_{jj}), \Omega = ((1-s) R + s I)^{-1}.
$$

From the simulation, the standardized differences between observed z score and the conditional mean has longer-tailed than N(0,1). We use a mixture of normals to model the heavier tail empirically. We model the conditional distribution as
$$
\hat{z}_j | \hat{z}_{-j} \sim \sum_{k=1}^{K} \pi_k N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},\frac{\sigma_k^2}{\Omega_{jj}}).
$$
The $\sigma_k$ is a fixed dense grid, the minimum value is $0.8$, the maximum value is $2\sqrt{max(\text{standardized difference}^2)}$. We estimate $\pi_k$ using `mixsqp`. The likelihood ratio test is based on the mixture distribution.

To check whether there is an allele switch issue, we test the hypothesis $H_0:$ the sign of $\hat{z}_j$ is correct, vs $H_1:$ the sign of $\hat{z}_j$ is flipped. The likelihood ratio is
$$
LR = \frac{p(-\hat{z}_j | \hat{z}_{-j})}{p(\hat{z}_j | \hat{z}_{-j})}
$$

There are several ways to estimate $s$. 

1. Let $R = U D U^\intercal$ be eigen-decomposition of $R$, $D$ has $p_1$ positive diagonal elements, $U = [U_1 \ U_2]$, $U_1$ is a $p$ by $p_1$ matrix, $U_2$ is a $p$ by $p_2=p-p_1$ matrix. We transform $\hat{z}$ into null space of $R$ by multiplying $U_2^\intercal$
$$
U_2^T \hat{z} \sim N(0, s I_{p_2}).
$$
Therefore, the MLE for $s$ using partial data is $\frac{1}{p_2} \hat{z}^\intercal U_2 U_2^\intercal \hat{z}$. If $s > 1$, we set it to 1.

2. We estimate $s$ by maximizing pseudo-likelihood, 
$$
l(s) = \sum_{j=1}^{p} \log p(\hat{z}_j | \hat{z}_{-j}), 0 \leq s \leq 1.
$$

3. We estimate $s$ by maximizing 
$$
\hat{z} \sim N(0, (1-s) R + s I), 0 \leq s \leq 1.
$$

**Simulation Setting**

In the following simulation, we extract 200 regions from UKB. We sample 10,000 UKB individuals to simulate z scores and generate LD in. We simulate data without signal. We sample 1,000 UKB individuals as reference panel. There are 1000 SNPs in each region.

```{r message=FALSE}
library(tibble)
library(dplyr)
out = readRDS('data/susierss_diagnostic_ukb_query.rds')
out = as_tibble(out)
out_nosignal_ukb_in_1 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'restricted')
out_nosignal_ukb_in_2 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'pseudo')
out_nosignal_ukb_in_3 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'likelihood')

out_nosignal_ukb_ref_1 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'restricted')
out_nosignal_ukb_ref_2 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'pseudo')
out_nosignal_ukb_ref_3 = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'likelihood')

out_nosignal_ukb_flip_in_1 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'restricted')
out_nosignal_ukb_flip_in_2 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'pseudo')
out_nosignal_ukb_flip_in_3 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample',
                                     diagnostic.method == 'likelihood')

out_nosignal_ukb_flip_ref_1 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'restricted')
out_nosignal_ukb_flip_ref_2 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'pseudo')
out_nosignal_ukb_flip_ref_3 = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref',
                                     diagnostic.method == 'likelihood')
```

## Simulation 1.1: no signal, no flip (z = 0)

### Check $s$

We compare the $s$ using different LD matrix.

```{r}
out_nosignal_ukb_in_1_s = apply(out_nosignal_ukb_in_1, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_in_2_s = apply(out_nosignal_ukb_in_2, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_in_3_s = apply(out_nosignal_ukb_in_3, 1, function(x) x$diagnostic.res$s)

out_nosignal_ukb_ref_1_s = apply(out_nosignal_ukb_ref_1, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_ref_2_s = apply(out_nosignal_ukb_ref_2, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_ref_3_s = apply(out_nosignal_ukb_ref_3, 1, function(x) x$diagnostic.res$s)

par(mfrow=c(1,3))
plot(out_nosignal_ukb_in_1_s, out_nosignal_ukb_ref_1_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='MLE partial data')
abline(0,1)
plot(out_nosignal_ukb_in_2_s, out_nosignal_ukb_ref_2_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='Pseudo likelihood')
abline(0,1)
plot(out_nosignal_ukb_in_3_s, out_nosignal_ukb_ref_3_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='likelihood')
abline(0,1)
par(mfrow=c(1,2))
plot(out_nosignal_ukb_ref_1_s, out_nosignal_ukb_ref_2_s, xlab = 's from MLE partial data', ylab = 's from Pseudo likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
plot(out_nosignal_ukb_ref_2_s, out_nosignal_ukb_ref_3_s, xlab = 's from Pseudo likelihood', ylab = 's from likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
```

The estimated $s$ is larger when the LD matrix is from reference panel. Since the results from pseudo likelihood and likelihood are similar, we ignore s from pseudo likelihood below.

Speed
```{r}
sp = rbind(summary(out_nosignal_ukb_in_1$diagnostic.DSC_TIME),
           summary(out_nosignal_ukb_in_2$diagnostic.DSC_TIME),
           summary(out_nosignal_ukb_in_3$diagnostic.DSC_TIME))
rownames(sp) = c('MLEpartial', 'pseudo likelihood', 'likelihood')
sp
```


### Check conditional mean and Likelihood

```{r}
z = unlist(lapply(out_nosignal_ukb_in_1$diagnostic.res, function(x) x$post$z))
out_nosignal_ukb_in_1_z = unlist(lapply(out_nosignal_ukb_in_1$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_in_1_lr = unlist(lapply(out_nosignal_ukb_in_1$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_in_1_lrmix = unlist(lapply(out_nosignal_ukb_in_1$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_in_1_lrmix[is.infinite(out_nosignal_ukb_in_1_lrmix)] = -1500

out_nosignal_ukb_in_3_z = unlist(lapply(out_nosignal_ukb_in_3$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_in_3_lr = unlist(lapply(out_nosignal_ukb_in_3$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_in_3_lrmix = unlist(lapply(out_nosignal_ukb_in_3$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_in_3_lrmix[is.infinite(out_nosignal_ukb_in_3_lrmix)] = -1500

out_nosignal_ukb_ref_1_z = unlist(lapply(out_nosignal_ukb_ref_1$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_ref_1_lr = unlist(lapply(out_nosignal_ukb_ref_1$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_ref_1_lrmix = unlist(lapply(out_nosignal_ukb_ref_1$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_ref_1_lrmix[is.infinite(out_nosignal_ukb_ref_1_lrmix)] = -1500

out_nosignal_ukb_ref_3_z = unlist(lapply(out_nosignal_ukb_ref_3$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_ref_3_lr = unlist(lapply(out_nosignal_ukb_ref_3$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_ref_3_lrmix = unlist(lapply(out_nosignal_ukb_ref_3$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_ref_3_lrmix[is.infinite(out_nosignal_ukb_ref_3_lrmix)] = -1500
```

```{r}
par(mfrow=c(2,2))

qqnorm(out_nosignal_ukb_in_1_z, main='UKB z, LD in, s from MLE partial data')
qqline(out_nosignal_ukb_in_1_z)
qqnorm(out_nosignal_ukb_ref_1_z, main='UKB z, LD ref, s from MLE partial data')
qqline(out_nosignal_ukb_ref_1_z)

qqnorm(out_nosignal_ukb_in_3_z, main='UKB z, LD in, s from likelihood')
qqline(out_nosignal_ukb_in_3_z)
qqnorm(out_nosignal_ukb_ref_3_z, main='UKB z, LD ref, s from likelihood')
qqline(out_nosignal_ukb_ref_3_z)
```

We focus on those z scores with magnitude greater than 2.
```{r}
zlarge = which(abs(z) > 2)
par(mfrow=c(2,2))
plot(out_nosignal_ukb_in_1_lr[zlarge], out_nosignal_ukb_in_3_lr[zlarge], xlim = c(-50,10), ylim = c(-50,10),
     xlab = 'logLR using s from MLE partial data', ylab = 'logLR using s from likelihood', 
     main = 'UKB z, LD in')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)

plot(out_nosignal_ukb_ref_1_lr[zlarge], out_nosignal_ukb_ref_3_lr[zlarge], xlim = c(-50,50), ylim = c(-50,50),
     xlab = 'logLR using s from MLE partial data', ylab = 'logLR using s from likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)

plot(out_nosignal_ukb_in_1_lrmix[zlarge], out_nosignal_ukb_in_3_lrmix[zlarge], xlim = c(-50,10), ylim = c(-50,10),
     xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', 
     main = 'UKB z, LD in')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)

plot(out_nosignal_ukb_ref_1_lrmix[zlarge], out_nosignal_ukb_ref_3_lrmix[zlarge], xlim = c(-50,10), ylim = c(-50,10),
     xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)
```

## Simulation 1.2: no signal, flip one sign

Based on Simulation 1.1, we randomly flipped sign of one z score.

### Check $s$

```{r}
out_nosignal_ukb_flip_in_1_s = apply(out_nosignal_ukb_flip_in_1, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_flip_in_2_s = apply(out_nosignal_ukb_flip_in_2, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_flip_in_3_s = apply(out_nosignal_ukb_flip_in_3, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_flip_ref_1_s = apply(out_nosignal_ukb_flip_ref_1, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_flip_ref_2_s = apply(out_nosignal_ukb_flip_ref_2, 1, function(x) x$diagnostic.res$s)
out_nosignal_ukb_flip_ref_3_s = apply(out_nosignal_ukb_flip_ref_3, 1, function(x) x$diagnostic.res$s)

par(mfrow=c(1,3))
plot(out_nosignal_ukb_flip_in_1_s, out_nosignal_ukb_flip_ref_1_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='MLE partial data')
abline(0,1)
plot(out_nosignal_ukb_flip_in_2_s, out_nosignal_ukb_flip_ref_2_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='Pseudo likelihood')
abline(0,1)
plot(out_nosignal_ukb_flip_in_3_s, out_nosignal_ukb_flip_ref_3_s, xlab = 'UKB z, LD in', ylab='UKB z, LD ref', main='likelihood')
abline(0,1)
par(mfrow=c(1,2))
plot(out_nosignal_ukb_flip_ref_1_s, out_nosignal_ukb_flip_ref_2_s, xlab = 's from MLE partial data', ylab = 's from Pseudo likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
plot(out_nosignal_ukb_flip_ref_2_s, out_nosignal_ukb_flip_ref_3_s, xlab = 's from Pseudo likelihood', ylab = 's from likelihood', 
     main = 'UKB z, LD ref')
abline(0,1)
```

### Check conditional mean and Likelihood

```{r}
z_flip = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$z))

out_nosignal_ukb_flip_in_1_z = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_flip_in_1_l0 = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logl0))
out_nosignal_ukb_flip_in_1_l1 = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logl1))
out_nosignal_ukb_flip_in_1_lr = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_flip_in_1_l0mix = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_in_1_l1mix = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_in_1_lrmix = unlist(lapply(out_nosignal_ukb_flip_in_1$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_in_1_l1mix[is.infinite(out_nosignal_ukb_flip_in_1_l1mix)] = -1500
out_nosignal_ukb_flip_in_1_lrmix[is.infinite(out_nosignal_ukb_flip_in_1_lrmix)] = -1500

out_nosignal_ukb_flip_in_3_z = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_flip_in_3_l0 = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logl0))
out_nosignal_ukb_flip_in_3_l1 = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logl1))
out_nosignal_ukb_flip_in_3_lr = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_flip_in_3_l0mix = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_in_3_l1mix = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_in_3_lrmix = unlist(lapply(out_nosignal_ukb_flip_in_3$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_in_3_l1mix[is.infinite(out_nosignal_ukb_flip_in_3_l1mix)] = -1500
out_nosignal_ukb_flip_in_3_lrmix[is.infinite(out_nosignal_ukb_flip_in_3_lrmix)] = -1500

out_nosignal_ukb_flip_ref_1_z = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_flip_ref_1_l0 = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logl0))
out_nosignal_ukb_flip_ref_1_l1 = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logl1))
out_nosignal_ukb_flip_ref_1_lr = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_flip_ref_1_l0mix = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_ref_1_l1mix = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_ref_1_lrmix = unlist(lapply(out_nosignal_ukb_flip_ref_1$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_ref_1_l1mix[is.infinite(out_nosignal_ukb_flip_ref_1_l1mix)] = -1500
out_nosignal_ukb_flip_ref_1_lrmix[is.infinite(out_nosignal_ukb_flip_ref_1_lrmix)] = -1500

out_nosignal_ukb_flip_ref_3_z = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$post_z))
out_nosignal_ukb_flip_ref_3_l0 = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logl0))
out_nosignal_ukb_flip_ref_3_l1 = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logl1))
out_nosignal_ukb_flip_ref_3_lr = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logLR))
out_nosignal_ukb_flip_ref_3_l0mix = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_ref_3_l1mix = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_ref_3_lrmix = unlist(lapply(out_nosignal_ukb_flip_ref_3$diagnostic.res, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_ref_3_l1mix[is.infinite(out_nosignal_ukb_flip_ref_3_l1mix)] = -1500
out_nosignal_ukb_flip_ref_3_lrmix[is.infinite(out_nosignal_ukb_flip_ref_3_lrmix)] = -1500
```

The red points corresponding to z scores with flipped sign.

```{r}
idx = which(z != z_flip)
par(mfrow=c(1,2))
plot(out_nosignal_ukb_flip_in_1_lrmix, out_nosignal_ukb_flip_in_3_lrmix, xlim = c(-50,20), ylim = c(-50,20),
     xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', 
     main = 'UKB z, LD in')
points(out_nosignal_ukb_flip_in_1_lrmix[idx], out_nosignal_ukb_flip_in_3_lrmix[idx], pch=16, col='red')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)

plot(out_nosignal_ukb_flip_ref_1_lrmix, out_nosignal_ukb_flip_ref_3_lrmix, xlim = c(-50,20), ylim = c(-50,20),
     xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', 
     main = 'UKB z, LD ref')
points(out_nosignal_ukb_flip_ref_1_lrmix[idx], out_nosignal_ukb_flip_ref_3_lrmix[idx], pch=16, col='red')
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)
```

We focus on those z scores with magnitude greater than 2.

```{r}
zlarge = which(abs(z) > 2)
par(mfrow=c(1,2))
plot(out_nosignal_ukb_flip_in_1_lrmix[zlarge], out_nosignal_ukb_flip_in_3_lrmix[zlarge], main='UKB z, LD in', xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', xlim=c(-50,20), ylim=c(-50,20))
points(out_nosignal_ukb_flip_in_1_lrmix[intersect(idx, zlarge)], out_nosignal_ukb_flip_in_3_lrmix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)

plot(out_nosignal_ukb_flip_ref_1_lrmix[zlarge], out_nosignal_ukb_flip_ref_3_lrmix[zlarge], main='UKB z, LD ref', xlab = 'logLRmix using s from MLE partial data', ylab = 'logLRmix using s from likelihood', xlim=c(-50,20), ylim=c(-50,20))
points(out_nosignal_ukb_flip_ref_1_lrmix[intersect(idx, zlarge)], out_nosignal_ukb_flip_ref_3_lrmix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)
abline(h=0, lty=2)
abline(v=0, lty=2)
```

