---
title: "SuSiE RSS Diagnostic using UKB (simulation without signal, sigma2=1)"
author: "Yuxin Zou"
date: "3/14/2021"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

We want to

1. Check the consistency between z scores and LD matrix using $s^2$;

2. Detect allele switch issue (flipped sign in z score) using conditional distribution.

In diagnostic, we assume the following model,
$$
\hat{z} \sim N(Rz, R + s^2 I).
$$
Let $R = U D U^\intercal$ be eigen-decomposition of $R$, $D$ has $p_1$ positive diagonal elements, $U = [U_1 \ U_2]$, $U_1$ is a $p$ by $p_1$ matrix, $U_2$ is a $p$ by $p_2=p-p_1$ matrix. We transform $\hat{z}$ into null space of $R$ by multiplying $U_2^\intercal$
$$
U_2^T \hat{z} \sim N(0, s^2 I_{p_2}).
$$
Therefore, the MLE for $s^2$ using partial data is $\frac{1}{p_2} \hat{z}^\intercal U_2 U_2^\intercal \hat{z}$.

Under the null, the posterior distribution of $\hat{z}_j | \hat{z}_{-j}$ is:
$$
\hat{z}_j | \hat{z}_{-j} \sim N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},1/\Omega_{jj}), \Omega = (R + s^2 I)^{-1}.
$$

From the simulation, the standardized differences between observed z score and the conditional mean is longer-tailed than N(0,1). We use a mixture of normals to model the heavier tail empirically. We model the conditional distribution as
$$
\hat{z}_j | \hat{z}_{-j} \sim \sum_{k=1}^{K} \pi_k N(-\frac{1}{\Omega_{jj}} \Omega_{j,-j} \hat{z}_{-j},\frac{\sigma_k^2}{\Omega_{jj}}).
$$
The $\sigma_k$ is a fixed dense grid, the minimum value is $0.8$, the maximum value is $2\sqrt{max(\text{standardized difference}^2)}$. We estimate $\pi_k$ using `mixsqp`. The likelihood ratio test is based on the mixture distribution.

To check whether there is an allele switch issue, we test the hypothesis $H_0:$ the sign of $\hat{z}_j$ is correct, vs $H_1:$ the sign of $\hat{z}_j$ is flipped. The likelihood ratio is
$$
LR = \frac{p(-\hat{z}_j | \hat{z}_{-j})}{p(\hat{z}_j | \hat{z}_{-j})}
$$

**Simulation Setting**

In the following simulation, we extract 200 regions from UKB. We sample 10,000 UKB individuals to simulate z scores and generate in sample LD. We simulate data without signal. We sample 1,000 UKB individuals as reference panel. There are 1000 SNPs in each region.

```{r message=FALSE}
library(tibble)
library(dplyr)
out = readRDS('data/susierss_diagnostic_ukb_query.rds')
out = as_tibble(out)
out_nosignal_ukb_in = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample')
out_nosignal_ukb_ref = out %>% filter(simulate.n_signal == 0, flip_z.flip == FALSE, 
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref')

out_nosignal_ukb_flip_in = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'sample')
out_nosignal_ukb_flip_ref = out %>% filter(simulate.n_signal == 0, flip_z.flip == TRUE, flip_z.flip_pos == 'null',
                                     diagnostic.z_type == 'ukb', diagnostic.ld_type == 'ref')
```

## Simulation 1.1: no signal, no flip (z = 0)

### Check $s^2$

We compare the $s^2$ using different LD matrix.

```{r}
out_nosignal_ukb_in_s2 = apply(out_nosignal_ukb_in, 1, function(x) x$diagnostic.res1_sigma2$s2)
out_nosignal_ukb_ref_s2 = apply(out_nosignal_ukb_ref, 1, function(x) x$diagnostic.res1_sigma2$s2)

plot(out_nosignal_ukb_in_s2, out_nosignal_ukb_ref_s2, xlab = 'UKB z, in sample LD', ylab='UKB z, UKB ref LD')
abline(0,1)
```

The estimated $s^2$ is larger when the LD matrix is from reference panel.

### Check conditional mean and Likelihood

```{r}
z = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$z))
out_nosignal_ukb_in_mean = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$postmean))
out_nosignal_ukb_in_z = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$post_z))
out_nosignal_ukb_in_l0 = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logl0))
out_nosignal_ukb_in_l1 = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logl1))
out_nosignal_ukb_in_lr = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logLR))
out_nosignal_ukb_in_l0mix = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logl0mix))
out_nosignal_ukb_in_l1mix = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logl1mix))
out_nosignal_ukb_in_lrmix = unlist(lapply(out_nosignal_ukb_in$diagnostic.res1_sigma2, function(x) x$post$logLRmix))
out_nosignal_ukb_in_l1mix[is.infinite(out_nosignal_ukb_in_l1mix)] = -1500
out_nosignal_ukb_in_lrmix[is.infinite(out_nosignal_ukb_in_lrmix)] = -1500

out_nosignal_ukb_ref_mean = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$postmean))
out_nosignal_ukb_ref_z = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$post_z))
out_nosignal_ukb_ref_l0 = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logl0))
out_nosignal_ukb_ref_l1 = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logl1))
out_nosignal_ukb_ref_lr = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logLR))
out_nosignal_ukb_ref_l0mix = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logl0mix))
out_nosignal_ukb_ref_l1mix = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logl1mix))
out_nosignal_ukb_ref_lrmix = unlist(lapply(out_nosignal_ukb_ref$diagnostic.res1_sigma2, function(x) x$post$logLRmix))
out_nosignal_ukb_ref_l1mix[is.infinite(out_nosignal_ukb_ref_l1mix)] = -1500
out_nosignal_ukb_ref_lrmix[is.infinite(out_nosignal_ukb_ref_lrmix)] = -1500
```

```{r}
par(mfrow=c(3,2))
diff_in = z-out_nosignal_ukb_in_mean; diff_ref = z-out_nosignal_ukb_ref_mean;
ymax = ceiling(max(c(diff_in, diff_ref))); ymin = floor(min(c(diff_in, diff_ref)))
plot(diff_in, ylab='UKB z, in sample LD', main='z-conditional mean',cex=0.6, ylim = c(ymin, ymax))
plot(diff_ref, ylab='UKB z, UKB ref LD', ylim = c(ymin, ymax))

ymax = ceiling(max(c(out_nosignal_ukb_in_z, out_nosignal_ukb_ref_z))); ymin = floor(min(c(out_nosignal_ukb_in_z, out_nosignal_ukb_ref_z)))
plot(out_nosignal_ukb_in_z, ylab='UKB z, in sample LD', main='standardized difference',cex=0.6, ylim = c(ymin, ymax))
plot(out_nosignal_ukb_ref_z, ylab='UKB z, UKB ref LD', ylim = c(ymin, ymax))

qqnorm(out_nosignal_ukb_in_z, main='Standardized Difference in sample LD, UKB z')
qqline(out_nosignal_ukb_in_z)
qqnorm(out_nosignal_ukb_ref_z, main='Standardized Difference ref LD, UKB z')
qqline(out_nosignal_ukb_ref_z)

par(mfrow=c(2,2))
plot(out_nosignal_ukb_in_l0, out_nosignal_ukb_in_l1, main='UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative',
     ylim = c(2*min(out_nosignal_ukb_in_l0), max(out_nosignal_ukb_in_l1)))
abline(0,1)
plot(out_nosignal_ukb_ref_l0, out_nosignal_ukb_ref_l1, main='UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative',
     ylim = c(2*min(out_nosignal_ukb_ref_l0), max(out_nosignal_ukb_ref_l1)))
abline(0,1)

plot(out_nosignal_ukb_in_l0mix, out_nosignal_ukb_in_l1mix, main='Mixure model, UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_in_l0mix), max(out_nosignal_ukb_in_l1mix)))
abline(0,1)
plot(out_nosignal_ukb_ref_l0mix, out_nosignal_ukb_ref_l1mix, main='Mixure model, UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim = c(2*min(out_nosignal_ukb_ref_l0mix), max(out_nosignal_ukb_ref_l1mix)))
abline(0,1)
```

We focus on those z scores with magnitude greater than 2.
```{r}
zlarge = which(abs(z) > 2)
par(mfrow=c(2,2))
plot(out_nosignal_ukb_in_l0[zlarge], out_nosignal_ukb_in_l1[zlarge], main='UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative',
     ylim = c(2*min(out_nosignal_ukb_in_l0[zlarge]), max(out_nosignal_ukb_in_l1[zlarge])),
     xlim = c(-20,5))
abline(0,1)
plot(out_nosignal_ukb_ref_l0[zlarge], out_nosignal_ukb_ref_l1[zlarge], main='UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative',
     ylim = c(2*min(out_nosignal_ukb_ref_l0[zlarge]), max(out_nosignal_ukb_ref_l1[zlarge])),
     xlim = c(-20,5))
abline(0,1)

plot(out_nosignal_ukb_in_l0mix[zlarge], out_nosignal_ukb_in_l1mix[zlarge], main='Mixure model, UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_in_l0mix), max(out_nosignal_ukb_in_l1mix)), xlim = c(-20,5))
abline(0,1)
plot(out_nosignal_ukb_ref_l0mix[zlarge], out_nosignal_ukb_ref_l1mix[zlarge], main='Mixure model, UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim = c(2*min(out_nosignal_ukb_ref_l0mix), max(out_nosignal_ukb_ref_l1mix)), xlim = c(-20,5))
abline(0,1)
```

## Simulation 1.2: no signal, flip one sign

Based on Simulation 1.1, we randomly flipped sign of one z score.

### Check $s^2$

```{r}
out_nosignal_ukb_flip_in_s2 = apply(out_nosignal_ukb_flip_in, 1, function(x) x$diagnostic.res1_sigma2$s2)
out_nosignal_ukb_flip_ref_s2 = apply(out_nosignal_ukb_flip_ref, 1, function(x) x$diagnostic.res1_sigma2$s2)

plot(out_nosignal_ukb_flip_in_s2, out_nosignal_ukb_flip_ref_s2, xlab = 'UKB z, in sample LD', ylab='UKB z, UKB ref LD')
abline(0,1)
```

### Check conditional mean and Likelihood

```{r}
z_flip = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$z))
out_nosignal_ukb_flip_in_mean = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$postmean))
out_nosignal_ukb_flip_in_z = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$post_z))
out_nosignal_ukb_flip_in_l0 = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logl0))
out_nosignal_ukb_flip_in_l1 = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logl1))
out_nosignal_ukb_flip_in_lr = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logLR))
out_nosignal_ukb_flip_in_l0mix = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_in_l1mix = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_in_lrmix = unlist(lapply(out_nosignal_ukb_flip_in$diagnostic.res1_sigma2, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_in_l1mix[is.infinite(out_nosignal_ukb_flip_in_l1mix)] = -1500
out_nosignal_ukb_flip_in_lrmix[is.infinite(out_nosignal_ukb_flip_in_lrmix)] = -1500

out_nosignal_ukb_flip_ref_mean = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$postmean))
out_nosignal_ukb_flip_ref_z = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$post_z))
out_nosignal_ukb_flip_ref_l0 = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logl0))
out_nosignal_ukb_flip_ref_l1 = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logl1))
out_nosignal_ukb_flip_ref_lr = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logLR))
out_nosignal_ukb_flip_ref_l0mix = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logl0mix))
out_nosignal_ukb_flip_ref_l1mix = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logl1mix))
out_nosignal_ukb_flip_ref_lrmix = unlist(lapply(out_nosignal_ukb_flip_ref$diagnostic.res1_sigma2, function(x) x$post$logLRmix))
out_nosignal_ukb_flip_ref_l1mix[is.infinite(out_nosignal_ukb_flip_ref_l1mix)] = -1500
out_nosignal_ukb_flip_ref_lrmix[is.infinite(out_nosignal_ukb_flip_ref_lrmix)] = -1500
```

The red points corresponding to z scores with flipped sign.

```{r}
par(mfrow=c(3,2))
idx = which(z != z_flip)
diff_in = z_flip-out_nosignal_ukb_flip_in_mean; diff_ref = z_flip-out_nosignal_ukb_flip_ref_mean;
ymax = ceiling(max(c(diff_in, diff_ref))); ymin = floor(min(c(diff_in, diff_ref)))
plot(diff_in, ylab='UKB z, in sample LD', main='z-conditional mean',cex=0.6, ylim = c(ymin, ymax))
points(idx, diff_in[idx], col='red', pch=16)
plot(diff_ref, ylab='UKB z, UKB ref LD', ylim = c(ymin, ymax))
points(idx, diff_ref[idx], col='red', pch=16)

ymax = ceiling(max(c(out_nosignal_ukb_flip_in_z, out_nosignal_ukb_flip_ref_z))); ymin = floor(min(c(out_nosignal_ukb_flip_in_z, out_nosignal_ukb_flip_ref_z)))
plot(out_nosignal_ukb_flip_in_z, ylab='UKB z, in sample LD', main='standardized difference',cex=0.6, ylim = c(ymin, ymax))
points(idx, out_nosignal_ukb_flip_in_z[idx], col='red', pch=16)
plot(out_nosignal_ukb_flip_ref_z, ylab='UKB z, UKB ref LD', ylim = c(ymin, ymax))
points(idx, out_nosignal_ukb_flip_ref_z[idx], col='red', pch=16)

qqnorm(out_nosignal_ukb_flip_in_z, ylab='Standardized Difference in sample LD, UKB z')
qqline(out_nosignal_ukb_flip_in_z)
qqnorm(out_nosignal_ukb_flip_ref_z, ylab='Standardized Difference ref LD, UKB z')
qqline(out_nosignal_ukb_flip_ref_z)

par(mfrow=c(2,2))
plot(out_nosignal_ukb_flip_in_l0, out_nosignal_ukb_flip_in_l1, main='UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_in_l0), max(out_nosignal_ukb_flip_in_l1)))
points(out_nosignal_ukb_flip_in_l0[idx], out_nosignal_ukb_flip_in_l1[idx], col='red', pch=16)
abline(0,1)

plot(out_nosignal_ukb_flip_ref_l0, out_nosignal_ukb_flip_ref_l1, main='UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_ref_l0), max(out_nosignal_ukb_flip_ref_l1)))
points(out_nosignal_ukb_flip_ref_l0[idx], out_nosignal_ukb_flip_ref_l1[idx], col='red', pch=16)
abline(0,1)


plot(out_nosignal_ukb_flip_in_l0mix, out_nosignal_ukb_flip_in_l1mix, main='Mixure model, UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_in_l0mix), max(out_nosignal_ukb_flip_in_l1mix)))
points(out_nosignal_ukb_flip_in_l0mix[idx], out_nosignal_ukb_flip_in_l1mix[idx], col='red', pch=16)
abline(0,1)

plot(out_nosignal_ukb_flip_ref_l0mix, out_nosignal_ukb_flip_ref_l1mix, main='Mixure model, UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_ref_l0mix), max(out_nosignal_ukb_flip_ref_l1mix)))
points(out_nosignal_ukb_flip_ref_l0mix[idx], out_nosignal_ukb_flip_ref_l1mix[idx], col='red', pch=16)
abline(0,1)
```

We focus on those z scores with magnitude greater than 2.

```{r}
par(mfrow=c(2,2))
plot(out_nosignal_ukb_flip_in_l0[zlarge], out_nosignal_ukb_flip_in_l1[zlarge], main='UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_in_l0[zlarge]), max(out_nosignal_ukb_flip_in_l1[zlarge])))
points(out_nosignal_ukb_flip_in_l0[intersect(idx, zlarge)], out_nosignal_ukb_flip_in_l1[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)

plot(out_nosignal_ukb_flip_ref_l0[zlarge], out_nosignal_ukb_flip_ref_l1[zlarge], main='UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_ref_l0[zlarge]), max(out_nosignal_ukb_flip_ref_l1[zlarge])))
points(out_nosignal_ukb_flip_ref_l0[intersect(idx, zlarge)], out_nosignal_ukb_flip_ref_l1[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)

plot(out_nosignal_ukb_flip_in_l0mix[zlarge], out_nosignal_ukb_flip_in_l1mix[zlarge], main='Mixure model, UKB z, in sample LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_in_l0mix[zlarge]), max(out_nosignal_ukb_flip_in_l1mix[zlarge])))
points(out_nosignal_ukb_flip_in_l0mix[intersect(idx, zlarge)], out_nosignal_ukb_flip_in_l1mix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)

plot(out_nosignal_ukb_flip_ref_l0mix[zlarge], out_nosignal_ukb_flip_ref_l1mix[zlarge], main='Mixure model, UKB z, UKB ref LD', xlab='log likelihood under null', ylab='log likelihood under alternative', ylim=c(2*min(out_nosignal_ukb_flip_ref_l0mix[zlarge]), max(out_nosignal_ukb_flip_ref_l1mix[zlarge])))
points(out_nosignal_ukb_flip_ref_l0mix[intersect(idx, zlarge)], out_nosignal_ukb_flip_ref_l1mix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)
```

Comparing LR from Simulation 1.1 and 1.2,
```{r}
par(mfrow=c(1,2))
plot(out_nosignal_ukb_in_lrmix[zlarge], out_nosignal_ukb_flip_in_lrmix[zlarge],
     xlab='logLRmix from Simu1.1',
     ylab='logLRmix from Simu1.2', main='Mixure model, UKB z, in sample LD', ylim=c(-10,15), xlim=c(-1500, 10))
points(out_nosignal_ukb_in_lrmix[intersect(idx, zlarge)], out_nosignal_ukb_flip_in_lrmix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)
abline(h=0,v=0,lty=4)
plot(out_nosignal_ukb_ref_lrmix[zlarge], out_nosignal_ukb_flip_ref_lrmix[zlarge],
     xlab='logLRmix from Simu1.1',
     ylab='logLRmix from Simu1.2', main='Mixure model, UKB z, UKB ref LD', ylim=c(-10,15), xlim=c(-1500, 10))
points(out_nosignal_ukb_ref_lrmix[intersect(idx, zlarge)], out_nosignal_ukb_flip_ref_lrmix[intersect(idx, zlarge)], col='red', pch=16)
abline(0,1)
abline(h=0,v=0,lty=4)
```

